\begin{titlepage}
  \centering
  \vspace*{2cm}
  {\heiti\zihao{2} 面向 LangGraph 智能体的 AI 原生应用层故障注入机制设计与实现\par}
  \vspace{2cm}
  {\zihao{3} 复旦大学计算与智能创新学院\par}
  \vspace{0.8cm}
  {\zihao{3} 本科毕业论文\par}
  \vfill
  \begin{tabular}{rl}
    学生姓名： & 陈董祺 \\
    学号： & 22302010052  \\
    专业： & 软件工程 \\
    指导教师： & 彭鑫 \\
    完成日期： & \today
  \end{tabular}
  \vspace*{2cm}
\end{titlepage}

\chapter*{摘\quad 要}
\addcontentsline{toc}{chapter}{摘要}

随着大语言模型和智能体编排框架进入真实业务系统，AI 原生应用已经不再是单次模型调用，而是由模型、Agent 状态机、工具调用、记忆系统、跨服务通信和传统后端共同构成的复杂软件系统。传统混沌工程主要面向主机、容器、网络和进程等资源层对象，AgentOps 平台则更侧重 Trace 采集、Bad Case 归档和提示词迭代。二者都难以稳定复现“某个工具调用高延迟”、“中间状态被清空”、“远程子 Agent 不可达”、“模型输出结构被破坏”等 AI 原生应用层故障。因此，如何在不大规模改写业务逻辑的前提下，建立可配置、可观测、可复现的应用层故障注入机制，是 AI 原生软件可靠性验证中的一个现实问题。

本文以 TripSphere 智能旅行系统为研究对象，结合其前端交互、LangGraph 编排、工具调用、远程 Order Agent、Java gRPC 微服务和 OpenTelemetry 可观测组件，分析 AI 原生应用的层次化结构，并抽象出三条主要链路：REST 行程规划链、CopilotKit ReAct 对话链、Chat 到 A2A 再到 Order 的跨服务链。围绕这些链路，本文提出面向 LangGraph 智能体的应用层故障模型，将故障划分为工具/RPC 延迟与异常、gRPC 错误码、工具响应篡改、状态清空、路由扰动、远程 Agent 丢失和 Trace 断链等类型；设计了基于实验标识、请求头 DSL、运行时注册表、上下文传播和 OpenTelemetry Span 属性的低侵入注入机制；并在 TripSphere 三条链路中落地请求级注入、命中证据记录和实验复跑流程。

本文的主要贡献包括：第一，从系统视角给出 TripSphere AI 原生应用的层次化架构分析，强调 Agent、工具、状态、远程服务与底层微服务之间的故障传播关系；第二，设计并整理一套面向应用层的可复现故障注入机制，使单次请求、单条 Trace 和单个工具调用可以被精确注入并留下可审计证据；第三，完成覆盖三条链路的四象限实验，使用 Locust、Tempo 和质量记录串联故障声明、注入命中、HTTP 指标和业务质量指标。实验结果表明，AI 原生应用中存在 HTTP 不可见退化、fail-fast 延迟反转和跨 Agent 故障放大等现象，仅依赖响应码和端到端延迟无法判断真实业务质量。

\noindent\textbf{关键词：}AI 原生应用，LangGraph，故障注入，混沌工程，可观测性

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

As large language models and agent orchestration frameworks move into real software systems, AI-native applications are no longer isolated model calls, but complex software systems composed of models, agent state machines, tool invocations, memory systems, cross-service communication, and conventional backend services. Traditional chaos engineering primarily targets resource-layer objects such as hosts, containers, networks, and processes, while AgentOps platforms focus more on trace collection, bad-case archiving, and prompt iteration. Neither is well-suited to reproducibly triggering application-layer failures such as high-latency tool calls, cleared intermediate state, unreachable remote sub-agents, or corrupted model outputs. Therefore, how to establish a configurable, observable, and reproducible application-layer fault-injection mechanism without extensively rewriting business logic is a practical problem in AI-native software reliability verification.

Using the TripSphere intelligent travel system as the target application, this thesis analyzes a layered AI-native architecture composed of frontend interaction, LangGraph orchestration, tool invocation, a remote Order Agent, Java gRPC microservices, and OpenTelemetry observability components. It abstracts three main execution chains: the REST itinerary-planning chain, the CopilotKit ReAct conversation chain, and the cross-service Chat-to-A2A-to-Order chain. Based on these chains, the thesis proposes an application-layer fault model for LangGraph agents, classifying faults into tool/RPC latency and exceptions, gRPC error codes, tool-response mutation, state clearing, route perturbation, remote-agent loss, and trace breakage; designs a low-intrusion injection mechanism based on experiment identifiers, a request-header DSL, a runtime registry, context propagation, and OpenTelemetry span attributes; and implements request-level injection, hit evidence recording, and experiment replay across the three TripSphere chains.

The main contributions are threefold. First, the thesis provides a system-level hierarchical architecture analysis of the TripSphere AI-native application, emphasizing how failures propagate among agents, tools, state, remote services, and underlying microservices. Second, it designs and organizes a reproducible application-layer fault-injection mechanism that allows a single request, trace, or tool invocation to be precisely injected while leaving auditable evidence. Third, it completes four-quadrant experiments over the three chains, using Locust, Tempo, and quality records to connect fault declarations, injection hits, HTTP metrics, and business-quality metrics. The results show that AI-native applications may suffer from HTTP-invisible degradation, fail-fast latency inversion, and cross-agent fault amplification, so response codes and end-to-end latency alone cannot represent real business quality.

\noindent\textbf{Keywords:} AI-native application, LangGraph, fault injection, chaos engineering, observability
